\color{navyblue} Difference-in-Differences Estimation with Spatial Spillovers

### 2021-05-03 latex/paper/Spillover.tex
5-3 Draft

\usepackage{dcolumn}
\newcolumntype{d}[1]{D..{#1}}
\newcommand{\mcc}[1]{\multicolumn{1}{c@{}}{#1}}
\newcommand{\mcl}[1]{\multicolumn{1}{l@{}}{#1}}
\newcommand{\mcr}[1]{\multicolumn{1}{r@{}}{#1}}

@@

    \begin{abstract}
        Empirical work often uses treatment variables defined by geographic boundaries. When researchers ignore the common problem that the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I decompose this bias in two parts (1) the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment and (2) changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias as well as semi-parametrically estimate the spillover effects themselves. I extend this estimation strategy in an event-study framework following \citet{Callaway_SantAnna_2020} to allow for staggered treatment adoption. To highlight the importance of spillover effects, I revisit the analysis of three place-based policies and show the importance of directly controlling for general equilibrium spillovers caused by place-based policies.: the Tennessee Valley Authority \citep{Kline_Moretti_2014}, Federal Empowerment Zones \citep{Neumark_Young_2019}, and Community Health Centers \citep{Bailey_Goodman_Bacon_2015}. 
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Empirical work often considers settings where treatment is assigned to groups of units by geographic boundaries, but the effect of these treatments spillover onto `nearby' units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} In this setting, there are two types of spillover effects. First, there can be spillover effects onto control units. For example, individuals in control areas can travel to the treated areas and receive treatment (e.g. a hospital opening serves nearby residents) or changes in the economy in a treated area can have feedback loops that affect nearby economies (e.g. a new factory opening increases service sector spending). Second, there can be spillover effects onto other treated units. For example, forces of agglomeration can increase effects as treatment concentration increases (e.g. knowledge-sharing between workers), forces of congestion or market competition can decrease the effects of treatment (e.g. bidding up of wages cause employment effects of factory openings to be mitigated), and treatment effectiveness may be improved by multiple jurisidictions learning from one another.

@@

There is a nascent literature exploring estimation of direct and spillover effects which does not require a completely untreated cluster by using a potential outcomes framework. \citet{Vazquez-Bare_2019} presents a potential outcomes model that explicitly accounts for ``within-group'' spillovers in experiments which allows for seperate estimation of direct effect and spillover effects by a simple differences-in-means. I extend this work by focusing on difference-in-differences estimation in non-experimental settings.  \citet{Sävje_Aronow_Hudgens_2019} consider estimation of treatment effects by difference-in-differences, but they define treatment effects as the combination of direct effects and spillover effects. My paper develops a strategy to allow researchers to seperately identify treatment effects and spillovers. These estimates can be combined to estimate their definition of average treatment effect. The further advantage is that `net' treatment effects can be estimated at different levels of exposure which is more relevant for future policy decisions.

@@

My definition of the direct effect of treatment differs from \citet{Sävje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \expec{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \ \vert \ D_i = 1},
\] 
where the expectation is over individuals and over different levels of exposures. The difference between their treatment effect and mine becomes more clear by adding and subtracting terms and rearranging:
\begin{align*}
    &Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0) + Y_{i,t}(1, 0) - Y_{i,t}(0,0) + Y_{i,t}(0,0) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = \underbrace{Y_{i,t}(1, 0) - Y_{i,t}(0,0)}_{\text{Direct Effect}} + \underbrace{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0)}_{\text{Spillover on Treated}} - \underbrace{(Y_{i,t}(0, h(\vec{D}, i)) - Y_{i,t}(0,0))}_{\text{Spillover on Control}} 
\end{align*}
The individual treatment effect \citet{Sävje_Aronow_Hudgens_2019} use is the direct effect of treatment plus the difference between spillover on treated units and spillover on control units.

I prefer my definition of the direct effect because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. The alternative estimand averages over exposure intensities and therefore is difficult to be used by a policy maker in a particular location. That is, a county deciding whether to implement a policy will want to consider the treatment effect plus the difference in spillover effects at \textit{a particular} exposure intensity. The estimand proposed by \citet{Sävje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at specific exposure intensities while my preferred method of identifying the three effects seperately can. Estimates of the direct effect and spillover effect can be later combined to estimate the average treatment effect as defined by \citet{Sävje_Aronow_Hudgens_2019}.

@@

% ------------------------------------------------------------------------------
\section{Estimation of Treatment Effects with Spillovers}
\label{sec:estimation}
% ------------------------------------------------------------------------------

@@

Under mild conditions, it is possible to control directly in a regression for the two sources of spillover effects and to have unbiased estimates of the direct effect of treatment. A pair of indicators that contains all control/treated units respectively will estimate the average spillover effect on control/treated units and remove these sources of bias. The assumptions required to do this is that the researcher needs to be able to identify the maximum distance for which units are affected by spillovers and that there be unexposed units remaining.

@@

The intuition behind this is that the indicator interacted with treatment status serve as unbiased estimates of the \textit{average} spillover effect onto treated and control units and therefore remove these terms from the direct effect estimate, $\hat{\tau}$. The proof uses standard Frisch-Waugh-Lovell style logic on a set of indicator variables and the fact that if $S_{i} = 0$ then $h(\vec{D}, i) = \vec{0}$.\footnote{The derivation follows closely the work of \citet{Clarke_2019}.} This proposition shows that a researcher does not need to know how the spillovers occur over space in order to have an unbiased estimate for treatment effects. An indicator for being `near' treatment ineracted with treatment status is enough to have an unbiased estimate for the treatment effect.\footnote{In contexts with different measures of `distance', note that this result only requires an indicator that captures all affected units. For example, thinking about spillovers between industries, this requires the researcher to correctly identify which industries are affected by spillovers and which are not.}

@@

An improvement on this is to use a set of concentric ring indicators instead of a single large ring indicator to more flexibly estimate the spillover effects (e.g 0-20 miles, 20-40 miles and 40-60 miles from the nearest treated unit). Since the set of rings is collinear with a single large ring, the direct effect estimate retains its unbiased properties, while spillover effects can potentially be estimated more accurately. A simple example of how rings improves estimation of spillover effects can be seen in Figure \ref{fig:example_rings}.

@@

The figure displays observations of $\bar{Y}_{i,Post} - \bar{Y}_{i,Pre}$ on the y-axis and the distance to the nearest treated unit on the x-axis and treated units are placed at a zero distance for visualization purposes. The example has a counterfactual trend of 2 so any difference from 2 is due to either treatment/spillover effects and the error term. Panel (a) uses a single indicator for being between 0 and 60 miles from treatment and the difference between the bar and the estimated counterfactual trend is the estimated spillover effect which in this case is about 0.6. This number masks that units within the first 15 miles have a treatment effect of 1 or larger and units between 40 and 60 miles have 0 spillover effects.

Figure (b) on the other hand uses 4 indicators for every 15 mile interval (0-15, 15-30, 30-45, and 45-60 miles). This improves on the estimate of a single ring in two ways. The multiple rings is able to better capture the way spillover effects evolve over distance. Further, the ring that is included past the true maximum distance where spillover effects occur has a point estimate near zero. The rings method acts a semi-parametric estimator in the sense that as the number of rings increase and their width shrinks, the estimates would trace out the spillover effect function (subject to their being units outside of the rings). However, precision of estimates would decrease as the number of rings increase and therefore there is a bias-variance trade-off in the number and width of rings. \citet{Clarke_2017} proposes a cross-validation technique for optimally choosing the number of rings and their techniques that uses the out-of-sample mean square prediction error as a way to estimate the bias-variance tradeoff. Interested readers can see the paper for technical and implementation details.

@@

Now, I turn to analyzing how the specifications perform at estimating the spillover effects themselves. For the next table, I will predict their spillover value for each control unit from the point estimates on $\tilde{h}(\vec{D}, i)$. Then as a measure for how well spillovers are estimated, I will calculate one minus the mean squared prediction error normalized by the sum of squared spillovers (for comparison across data-generating processes), \[ 
    1 - \frac{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i) - \hat{\beta}_{\text{spill, control}} \tilde{h}(\vec{D}, i))^2}{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i))^2}.
\] This will produce a percentage of spillovers that are explained by the specification, so numbers closer to $100\%$ represent better modelling of the spillovers.

@@

% ------------------------------------------------------------------------------
\subsection{Guide for Researchers}
\label{sec:parameterize}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\section{Application in Place-Based Policy Analysis}
\label{sec:tva}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\section{Event Study with Spillovers}
\label{sec:event_study}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\section{Application on Community Health Centers}\label{sec:chc}
% ------------------------------------------------------------------------------

@@

I use the set of counties in the contiguous United States and the data generating process used is defined by the exposure mapping given in (\ref{eq:example_exposure}) with cutoff distance $\bar{d} = 40$ miles and the data generating process given in (\ref{eq:example_po}). A unit of observation is a US county and the periods are $t \in \{1, \dots, 20\}$ with treatment turning on after period $t = 10$. The unit and time fixed effects are generated by $\mu_t \sim N(0.2t, 0.1^2)$ and $\mu_i \sim N(6, 2^2)$ respectively, and the error term is $\varepsilon \sim N(0, 2^2)$. Last, the size of treatment and spillovers are as follows: $\beta_{\text{direct}} = 2$, $\beta_{\text{spill, control}} = 1$ and $\beta_{\text{spill, treat}} = 0$.

>>>>

@@

    \begin{abstract}
        Empirical work often uses treatment assigned following geographic boundaries. When the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I introduce a potential outcomes framework to model spillover effects and decompose the estimate's bias in two parts: (1) the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment and (2) changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias and semi-parametrically estimate the spillover effects themselves. I extend \citet{Callaway_SantAnna_2020} to allow for event-study estimates that control for spillovers. To highlight the importance of spillover effects, I revisit analyses of three place-based interventions. 
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Empirical work often considers settings where policy is targeted to groups of units by geographic boundaries, but the effect of these treatments spillover onto `nearby' units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} Place-based policies, programs that are targeted to specific locations, are known to have general-equilibrium effects that extend far past the area intended to receive treatment \citep{Kline_Moretti_2014b}. Spillover effects can occur on to both control and treated units. First, there can be spillover effects onto control units. For example, individuals in control areas can travel to the treated areas and receive treatment (e.g. a hospital opening serves nearby residents) or changes in the economy in a treated area can have feedback loops that affect nearby economies (e.g. a new factory opening increases service sector spending). Second, there can be spillover effects onto other treated units. For example, forces of agglomeration can increase effects as treatment concentration increases (e.g. knowledge-sharing between workers), forces of congestion or market competition can decrease the effects of treatment (e.g. bidding up of wages cause employment effects of factory openings to be mitigated), and treatment effectiveness may be improved by multiple jurisidictions learning from one another.

@@

There is a nascent literature exploring estimation of direct and spillover effects which does not require a completely untreated cluster by using a potential outcomes framework. \citet{Vazquez-Bare_2019} presents a potential outcomes model that explicitly accounts for ``within-group'' spillovers in experiments which allows for seperate estimation of direct effect and spillover effects by a simple differences-in-means. I extend this work by focusing on difference-in-differences estimation in non-experimental settings.  \citet{Savje_Aronow_Hudgens_2019} consider estimation of treatment effects by difference-in-differences, but they define treatment effects as the combination of direct effects and spillover effects. My paper develops a strategy to allow researchers to seperately identify treatment effects and spillovers. These estimates can be combined to estimate their definition of average treatment effect. The further advantage is that `net' treatment effects can be estimated at different levels of exposure which is more relevant for future policy decisions.

@@

My definition of the direct effect of treatment differs from \citet{Savje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \expec{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \ \vert \ D_i = 1},
\] 
where the expectation is over individuals and over different levels of exposures. The difference between their treatment effect and mine becomes more clear by adding and subtracting terms and rearranging:
\begin{align*}
    &Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0) + Y_{i,t}(1, 0) - Y_{i,t}(0,0) + Y_{i,t}(0,0) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = \underbrace{Y_{i,t}(1, 0) - Y_{i,t}(0,0)}_{\text{Direct Effect}} + \underbrace{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0)}_{\text{Spillover on Treated}} - \underbrace{(Y_{i,t}(0, h(\vec{D}, i)) - Y_{i,t}(0,0))}_{\text{Spillover on Control}} 
\end{align*}
The individual treatment effect \citet{Savje_Aronow_Hudgens_2019} use is the direct effect of treatment plus the difference between spillover on treated units and spillover on control units.

I prefer my definition of the direct effect because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. The alternative estimand averages over exposure intensities and therefore is difficult to be used by a policy maker in a particular location. That is, a county deciding whether to implement a policy will want to consider the treatment effect plus the difference in spillover effects at \textit{a particular} exposure intensity. The estimand proposed by \citet{Savje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at specific exposure intensities while my preferred method of identifying the three effects seperately can. Estimates of the direct effect and spillover effect can be later combined to estimate the average treatment effect as defined by \citet{Savje_Aronow_Hudgens_2019}.

@@

% ------------------------------------------------------------------------------
\section{Estimating $2\times 2$ Difference-in-Differences with Spillovers}
\label{sec:estimation}
% ------------------------------------------------------------------------------

@@

Under mild conditions, it is possible to control directly in a regression for the two sources of spillover effects and unbiasedly estimate the direct effect of treatment. A pair of indicators that contains all control/treated units respectively will estimate the average spillover effect on control/treated units and remove these sources of bias. The assumptions required to do this is that the researcher needs to be able to identify the maximum distance for which units are affected by spillovers and that there be unexposed units remaining.

@@

The intuition behind this is that the indicator interacted with treatment status serve as unbiased estimates of the \textit{average} spillover effect onto treated and control units and therefore remove these terms from the direct effect estimate, $\hat{\tau}$.\footnote{The proof uses standard Frisch-Waugh-Lovell style logic on a set of indicator variables and the fact that if $S_{i} = 0$ then $h(\vec{D}, i) = \vec{0}$. The derivation follows closely the work of \citet{Clarke_2019}.} This proposition shows that a researcher does not need to know how the spillovers occur over space in order to have an unbiased estimate for treatment effects. An indicator for being `near' treatment ineracted with treatment status is enough to have an unbiased estimate for the treatment effect.\footnote{In contexts with different measures of `distance', note that this result only requires an indicator that captures all affected units. For example, thinking about spillovers between industries, this requires the researcher to correctly identify which industries are affected by spillovers and which are not.}

@@

@@

An improvement on this is to use a set of concentric ring indicators instead of a single large ring indicator to more flexibly estimate the spillover effects (e.g 0-20 miles, 20-40 miles and 40-60 miles from the nearest treated unit). Since the set of rings is collinear with a single large ring, the direct effect estimate retains its unbiased properties, while spillover effects can potentially be estimated more accurately. A simple example of how rings improves estimation of spillover effects can be seen in Figure \ref{fig:example_rings}. The figure displays observations of $\bar{Y}_{i,Post} - \bar{Y}_{i,Pre}$ on the y-axis and the distance to the nearest treated unit on the x-axis and treated units are placed at a zero distance for visualization purposes. The example has a counterfactual trend of 2 so any difference from 2 is due to either treatment/spillover effects and the error term. Panel (a) uses a single indicator for being between 0 and 60 miles from treatment and the difference between the bar and the estimated counterfactual trend is the estimated spillover effect which in this case is about 0.6. This number masks that units within the first 15 miles have a treatment effect of 1 or larger and units between 40 and 60 miles have 0 spillover effects.

Figure (b) on the other hand uses 4 indicators for every 15 mile interval (0-15, 15-30, 30-45, and 45-60 miles). This improves on the estimate of a single ring in two ways. Multiple rings are better able to capture the evolution of spillover effects over distance. Furthermore, the ring that is included past the true maximum distance where spillover effects occur has a point estimate near zero. The rings method acts a semi-parametric estimator in the sense that as the number of rings increase and their width shrinks, the estimates would trace out the spillover effect function (subject to their being units outside of the rings). However, precision of estimates would decrease as the number of rings increase and therefore there is a bias-variance trade-off in the number and width of rings. \citet{Clarke_2017} proposes a cross-validation technique for optimally choosing the number of rings and their techniques that uses the out-of-sample mean square prediction error as a way to estimate the bias-variance tradeoff. Interested readers can see the paper for technical and implementation details.

@@

Now, I turn to analyzing how the specifications perform at estimating the spillover effects themselves. For the next table, I predict the spillover effect for each control unit from the point estimates on $\tilde{h}(\vec{D}, i)$. Then as a measure for how well spillovers are estimated, I will calculate one minus the mean squared prediction error normalized by the sum of squared spillovers (for comparison across data-generating processes), \[ 
    1 - \frac{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i) - \hat{\beta}_{\text{spill, control}} \tilde{h}(\vec{D}, i))^2}{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i))^2}.
\] This will produce a percentage of spillovers that are explained by the specification, so numbers closer to $100\%$ represent better modelling of the spillovers.

@@

% ------------------------------------------------------------------------------
\subsection{Guide for Applied Researchers}
\label{sec:parameterize}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\subsection{Application in Place-Based Policy Analysis}
\label{sec:tva}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\section{Estimating Event Study with Spillovers}
\label{sec:event_study}
% ------------------------------------------------------------------------------

@@

% ------------------------------------------------------------------------------
\subsection{Application on Community Health Centers}\label{sec:chc}
% ------------------------------------------------------------------------------

@@

I use the set of counties in the contiguous United States and the data generating process used is defined by the exposure mapping given in (\ref{eq:h_within}) with cutoff distance $\bar{d} = 40$ miles and the data generating process given in (\ref{eq:example_po}). A unit of observation is a US county and the periods are $t \in \{1, \dots, 20\}$ with treatment turning on after period $t = 10$. The unit and time fixed effects are generated by $\mu_t \sim N(0.2t, 0.1^2)$ and $\mu_i \sim N(6, 2^2)$ respectively, and the error term is $\varepsilon \sim N(0, 2^2)$. Last, the size of treatment and spillovers are as follows: $\beta_{\text{direct}} = 2$, $\beta_{\text{spill, control}} = 1$ and $\beta_{\text{spill, treat}} = 0$.
